---
title: "RDOTruckingCherynLindsay"
author: "Cheryn"
date: "2025-11-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr, quietly = TRUE) # for csv
library(readxl, quietly = TRUE)
```

# RDO Trucking Operations Optimization
## Background
RDO Trucking, a division of R.D. Offutt Company, manages fleet operations and driver scheduling to fulfill internal transportation requests across multiple locations. **The trucking team is called to transfer equipment between stores, deliver to or retrieve equipment from customers’ field locations and other third-party sites.** Drivers on the trucking team are not expected to physically load or unload equipment at either origination or destination sites. Drivers do perform safety inspections before departing locations to ensure the load meets trucking regulations.

The trucking team estimates that 8% of loads will incur a 2.0 hour delay at the point-of-origination; such delays add time to the driver’s work day. An estimated 1.5% of loads will incur a 4.0 hour delay at origination. These delays are necessary to resolve outstanding permitting issues, or address loading safety issues.

## Key Decision Variables
-	Driver-to-work-order assignments
-	Route selection and sequencing
-	Load consolidation (combining multiple work orders)
-	Weekly schedule optimization

## Objective Function
Minimize total operating costs given the following constraints:

### Drivers’ Federal Hours-of-Service Regulations
-	Maximum 14 hours on-duty per day
-	Maximum 11 hours driving per day (12 in Texas)
-	Minimum 10-hour rest period between shifts              
-	Maximum 70 hours driving in 8 consecutive days (34-hour rest required to reset 70-hour clock) **DOES THIS ONE NOT INCLUDE THE REST FROM SUNDAY?**
	
### Operational Requirements
-	Drivers have a maximum number of overnight stays per week.
-	Work week: Monday 12:00 AM - Sunday 11:59 PM (local time)
-	Work on Sundays is not authorized 
-	Maximum of 4 hours of work on Saturdays
-	Vehicle gross weight limits must be observed

### Time Calculations
-	Estimated drivers’ daily driving hours is total planned distance divided by 50 (assume average daily travel rate of 50 MPH)
-	Total work time = Drive time + 0.5 hours + 0.5 hours per segment (origination -> destination)
-	Labor costs (including California overtime: 1.5× after 8 hours, 2× after 10 hours; all other states: 1.5× after 40 weekly hours) **WHAT ARE DRIVERS PAID?**
-	Fuel and maintenance costs proportional to distance and weight
-	Empty miles (deadhead) penalty
-	Driver preference violations (overnight stays)
-	Loads are delivered by their expected delivery date.
-	$175 meal and stay cost per night
-	0.5 hour for refuel

### Stretch Goals
-	Available yard space to leave order

```{r}
drivers_csv <- read_csv("Drivers.csv")
locations_csv <- read_csv("Locations.csv")
mileage_matrix <- read_xlsx("MileageMatrix.xlsx")
trucking_workorders_csv <- read_csv("TruckingWorkorders.csv")


week <- c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday")



```
```{r}
drivers_csv
```

